Immutable Source-Mapped Abstract Syntax Tree:\ \, \\ A Design Pattern for Refactoring Engine APIs

### 2016-08-11 paper.tex
Short section talking about the example plugin.

% Title portion
\title{Immutable Source-Mapped Abstract Syntax Tree:\ \\
A Design Pattern for Refactoring Engine APIs} % \\ \today}
\author{JEFFREY L.\ OVERBEY \and CARL R.\ WORLEY \affil{Auburn University}}

@@

\acmformat{Overbey, J. and Worley, C. 2013. Immutable Source-Mapped Abstract 
Syntax Tree.}

@@

There are two main variants to the Abstract Syntax Tree pattern for refactoring:
mutable and immutable. A mutable tree has nodes that can be easily added, 
changed, or deleted from the tree. These trees are often used internally by
compilers. For example, a compiler might determine that
\ttt{a~=~b*1} can be simplified to \ttt{a~=~b}, and then change the AST 
accordingly.

@@

    \item \textit{Provide an API that allows clients to specify what changes
        they want to make in terms of AST nodes.}  This API can then translate these into string edits.  For example, an API might offer a \textit{Rewriter} 
class. The client would inform the \textit{Rewriter} that it wants to 
effectively delete a particular \textit{FieldDeclaration} node from the AST:
\textit{rewriter.delete(field)}.  When it is ready to manipulate the source
code, it delegates this task to the rewriter:
\textit{rewriter.rewrite(sourceCode)}.  This is illustrated in
Figure~\ref{fig:seq-diagram}.  The Eclipse Java and C/C++ Development
Tools~\cite{cdt-refactoring} take this approach (with slightly different class
and method names).

@@

The topic of refactoring preprocessed code will not be treated in detail here,
but suffice it to say that a compiler's AST will almost certainly be inadequate.
For more details, see Overbey et al.~\cite{wrt2009} and 
Garrido~\cite{garrido05program}.

@@

%\section{Related Patterns and Guidelines}

>>>>

% Title portion
\title{Immutable Source-Mapped Abstract Syntax Tree:\ \, \\
 A Design Pattern for Refactoring Engine APIs} % \\ \today}
\author{JEFFREY L.\ OVERBEY \and CARL R.\ WORLEY \affil{Auburn University}}

@@

\acmformat{Overbey, J. and Worley, C. 2016. Immutable Source-Mapped Abstract 
Syntax Tree.}

@@

There are two main variants of abstract syntax tree:
mutable and immutable. A mutable tree has nodes that can be easily added, 
changed, or deleted from the tree. These trees are often used internally by
compilers. For example, a compiler might determine that
\ttt{a~=~b*1} can be simplified to \ttt{a~=~b}, and then change the AST 
accordingly.

@@

    \item \textit{Provide an API that allows clients to specify what changes
        they want to make in terms of AST nodes.}  This API can then translate these into string edits.  For example, an API might offer a \textit{Rewriter} 
class. The client would inform the \textit{Rewriter} that it wants to 
effectively delete a particular \textit{FieldDeclaration} node from the AST:
\textit{rewriter.delete(field)}.  When it is ready to manipulate the source
code, it delegates this task to the rewriter:
\textit{rewriter.rewrite(sourceCode)}.  This allows multiple edits to be made 
without conflicting with each other. This is illustrated in
Figure~\ref{fig:seq-diagram}.  The Eclipse Java and C/C++ Development
Tools~\cite{cdt-refactoring} take this approach (with slightly different class
and method names).

@@

The topic of refactoring preprocessed code will not be treated in detail here,
but suffice it to say that a compiler's AST will almost certainly be inadequate.
For more details, see Garrido~\cite{garrido2013} and Overbey et al.~\cite{wrt2009}.

@@

\section{Example}
We created a plugin for Eclipse that exhibits the \emph{Rewriter} variant of this
pattern. It performs the \emph{reverse conditional} refactoring described 
earlier in both JDT and CDT. The pattern is most visible in the plugin's
swapStatements method, Figure~\ref{fig:swapStatements}, which swaps the 
branches of the \emph{if} statement.

As can be seen, the AST provides simple methods to retrieve individual nodes as
objects. The desired changes are then described to the \emph{rewriter}, which
doesn't make any actual changes until all edits are collected. In the case of
this refactoring, the only other change is the negation of the conditional 
expression, which is performed in a similar manner, replacing the existing
expression with a negated one.

The pattern is similar in both CDT and JDT, to the extent that the plugin is
structured identically, with slightly different class names in each.

\begin{figure}
\begin{center}
\includegraphics[scale=0.30]{swapStatements.eps}
\caption{The \emph{swapStatements} method.\vspace*{-4em}}
\label{fig:swapStatements}
\end{center}
\end{figure}
